Study of mixing behaviour of spherical particles in fluid using stirrer

### 2024-03-01 manuscript/paper.tex
Add more examples and add content to paper

  \title{}
  \author[XXX]{Dinesh Adepu\corref{cor1}}
  \ead{d.dinesh@surrey.ac.uk}
  \author[University of Surrey]{Chuan Yu Wu}
  \ead{XXX}
\address[xxx]{xxx}

@@

\begin{abstract}
\end{abstract}

@@

\section{Governing equations}

\begin{equation}
  \label{eq:ce}
  \frac{d \rho}{d t} = - \rho \; \frac{\partial u_i}{\partial x_i},
\end{equation}
and conservation of linear momentum,
\begin{equation}
  \label{eq:me}
  \frac{d u_i}{d t} = \frac{1}{\rho} \; \frac{\partial \sigma_{ij}}{\partial x_j}
  + g_i,
\end{equation}
where $\rho$ is the density, $u_i$ is the $i$\textsuperscript{th} component of
the velocity field, $x_j$ is the $j$\textsuperscript{th} component of the
position vector, $g_i$ is the component of body force per unit mass and
$\sigma_{ij}$ is stress tensor.

The stress tensor is split into isotropic and deviatoric parts,
\begin{equation}
  \label{eq:stress_tensor_decomposition}
  \sigma_{ij} = - p \; \delta_{ij} + \sigma'_{ij},
\end{equation}
%
where $p$ is the pressure, $\delta_{ij}$ is the Kronecker delta function, and
$\sigma'_{ij}$ is the deviatoric stress.

The Jaumann's formulation for Hooke's stress provides us with the rate of
change of deviatoric stress,
\begin{equation}
  \label{eq:jaumann-stress-rate}
  \frac{d \sigma'_{ij}}{dt} = 2G (\dot{\epsilon}_{ij} - \frac{1}{3}
  \dot{\epsilon}_{kk} \delta_{ij}) + \sigma^{'}_{ik}  \Omega_{jk} +
  \Omega_{ik} \sigma^{'}_{kj},
\end{equation}
where $G$ is the shear modulus, $\dot{\epsilon}_{ij}$ is the strain rate tensor,
\begin{equation}
  \label{eq:strain-tensor}
  \dot{\epsilon}_{ij} = \frac{1}{2} \bigg(\frac{\partial u_i}{\partial x_j} +
  \frac{\partial u_j}{\partial x_i} \bigg),
\end{equation}
and $\Omega_{ij}$ is the rotation tensor,
\begin{equation}
  \label{eq:rotational-tensor}
  \Omega_{ij} = \frac{1}{2} \bigg(\frac{\partial u_i}{\partial x_j} -
  \frac{\partial u_j}{\partial x_i} \bigg).
\end{equation}

For a weakly-compressible or incompressible fluid, a viscous force is added:
\begin{equation}
  \label{eq:fluid-stress-decomposition}
  \sigma_{ij} = - p \delta_{ij} + 2 \eta \frac{\partial u_i}{\partial x_j}
\end{equation}
where $\eta$ is the kinematic viscosity of the fluid.

In both fluid and solid modelling the pressure is computed using an
isothermal equation of state, given as,
\begin{equation}
  \label{eq:pressure-equation}
  p = K \bigg(\frac{\rho}{\rho_{0}} - 1 \bigg),
\end{equation}
where $K = \rho_{0} c_0^2$ is the bulk modulus. Here, the constants $c_0$ and
$\rho_0$ are the reference speed of sound and density, respectively. For solids,
$c_0$ is computed as $\sqrt{\frac{E}{3 (1 - 2 \nu)\rho_{0}}}$, $\nu$ is the
Poisson ratio, $E$ is the Young's modulus.

\section{Smoothed Particle Hydrodynamics}
In the current section, we introduce the basics of smoothed particle
hydrodynamics. A fundamental overview of the discrete approximations of
function, derivative, and divergence is discussed.

\subsection{Function Approximation}
From the dirac delta function definition, the function value of a smooth
function $A$ defined over a domain $\Omega$ at point $x$ can be written as
\begin{equation}
  \label{eq:dirac_repr}
  A(\ten{x}) = \int_{\Omega}\> A(\ten{x}') \> \delta(\ten{x} - \ten{x}') \> d\ten{x}',
\end{equation}
where $\delta(\ten{x} - \ten{x}')$ is a Dirac-delta function with
\begin{equation}
  \label{eq:unity}
  \int_{\Omega}\> \delta(\ten{x}) \> d\ten{x} = 1.
\end{equation}
In SPH, the Dirac delta function is replaced with a compact, smooth function $W$,
chosen to be as an even function. The function value of $A$ at $\ten{x}$
interpolated with the smooth kernel $W$ is given by
\begin{equation}
  \label{eq:kernel_repr}
  A(\ten{x}) = \int_{\Omega}\> A(\ten{x}') \> W(\ten{x} - \ten{x}', h)  \> d\ten{x}' + O(h^2).
\end{equation}
Here, kernel $W$ satisfies the following conditions,
\begin{equation}
  \label{eq:unity}
  \int_{\Omega}\> W(\ten{x} - \ten{x}', h)  \> d\ten{x}' = 1.
\end{equation}

\begin{equation}
  \label{eq:kernel_delta}
  \lim_{h \to 0} W(\boldsymbol{x} - \boldsymbol{x}', h) = \delta(\boldsymbol{x} - \boldsymbol{x}', h).
\end{equation}
\begin{equation}
  \label{eq:compact_support}
  W(\boldsymbol{x} - \boldsymbol{x}', h) = 0 \>\>\>\>    \text{when}  \>\>\> |\boldsymbol{x} - \boldsymbol{x}'| > kh.
\end{equation}
Here k is a constant, which determines the the non-zero region $W$ for a point
at position vector $\boldsymbol{x}$. In the current work we have used Quintic
% spline in all our simulation cases. Quintic spline \parencite{Violeau16}, is
given as,
\begin{equation}
  \label{eq:quintic_spline}
  W(r, h) =
  \begin{cases}
    \sigma_5\left( (3-q)^5 - 6(2-q)^5 + 15(1-q)^5 \right), & \textrm{for} \ 0\leq q \leq 1,\\
    \sigma_5\left( (3-q)^5 - 6(2-q)^5 \right), & \textrm{for} \ 1 <  q \leq 2,\\
    \sigma_5 \ (3-q)^5 , & \textrm{for} \ 2 < q \leq 3,\\
    0, & \textrm{for} \ q>3,
  \end{cases}
\end{equation}
This type of interpolation leads to a 2nd order function accuracy
% \parencite{fatehi2011error}.

Assume that the domain is discretized into N particles with mass
$m_i$ and volume $\Delta v_i$, We have
\begin{equation}
  \label{eq:mass_repr}
  m_j = \Delta v_i \> \rho_i.
\end{equation}
Using \eqref{eq:mass_repr}, \eqref{eq:kernel_repr} is converted to
discrete form as follows
\begin{equation}
  \label{eq:discrete_form}
  A_i = \sum_{j \in \text{Neigh}(i)}\> \frac{m_j}{\rho_j} A_j\> W(\ten{x}_i - \ten{x}_j, h),
\end{equation}
where $A(\boldsymbol{x}_i) = A_i$ is the value of the field property of particle
i, similarly for $A_j$. Here, $\text{Neigh}(i)$ is the neighbours of particle $i$.

\subsection{Derivative Approximation}
% The discretized form of the derivative approximation~\parencite{Violeau16} is given
as,
\begin{equation}
  \label{intro:eq:derivative-discrete_form}
  \nabla A(\ten{x}_i) = \sum_j\> \frac{m_j}{\rho_j} A_j\> \nabla W(\ten{x}_i - \ten{x}_j, h).
\end{equation}

An alternative expression of the derivative approximation is derived by employing
density, i.e,
\begin{equation}
  \label{intro:eq:sph-continu-derivative-final}
  \nabla (A \rho^n) = n A \rho^{n - 1} \nabla \rho + \rho^n \nabla A,
\end{equation}
where $n$ is a real number. The new identity for the derivative of $A$ is
\begin{equation}
  \nabla A = \frac{1}{\rho^n} \left[ n A \rho^{n - 1} \nabla \rho + \rho^n \nabla A  \right],
\end{equation}
using \cref{eq:discrete_form,intro:eq:derivative-discrete_form}, the new
derivative approximation can be written as
\begin{equation}
  \label{intro:eq:derivative-discrete_form}
  \nabla A(\ten{x}_i) = \frac{1}{\rho^n(\ten{x}_i)} \sum\>
  m_j \left(A_j \rho^{n-1}(\ten{x}_j) - n A_i \rho^{n-1}(\ten{x}_i) \right)\> \nabla W(\ten{x}_i - \ten{x}_j, h).
\end{equation}
In the current work, we use a symmetric expression, with $n=-1$, resulting in
\begin{equation}
  \nabla A(\ten{x}_i) = \rho_i \sum_{j} m_j \left(\frac{A_i}{\rho_i^2} + \frac{A_j}{\rho_j^2}\right) \nabla W_{ij}.
\end{equation}
A symmetric gradient operator ensures equal and opposite forces acting on the
particles.

\subsection{Divergence Approximation}
We now look at an SPH approximation of the divergence operator. An approximation
of the divergence operator is used to discretize the \cref{eq:ce}. It is given
as,
\begin{equation}
  \label{intro:eq:discrete_derivative}
  \nabla \cdot \ten{v}_i = \sum\> \frac{m_j}{\rho_j} \ten{v}_j \cdot \nabla W(\ten{x}_i - \ten{x}_j, h).
\end{equation}
Similar to the derivative approximation, we utilize the following identity to approximate the divergence operator,
\begin{equation}
  \label{intro:eq:sph-continu-div-intermediate}
  \nabla \cdot (\ten{v} \rho^n) = \rho^n \nabla \cdot \ten{v} + n  \rho^{n - 1} \ten{v} \cdot \nabla \rho.
\end{equation}
In the current work, we employ expression for $n=1$. The discrete SPH approximation for $n=1$ is given as
\begin{equation}
  \label{intro:eq:sph-continu-div-final}
  \nabla \cdot \ten{v}(\ten{x}_i) = \frac{1}{\rho_i} \sum_{j} m_j \left(\ten{v}_j - \ten{v}_i\right) \cdot \nabla W_{ij}.
\end{equation}

\subsection{SPH Discretization}\label{sec:ctvf-sph-equations}

In the current work, both fluid and solid modelling uses the same continuity
and pressure evolution equation. The SPH discretization of the continuity
equation~\eqref{eq:ce-tvf} and the pressure evolution
equation~\eqref{eq:edac-p-evolve} respectively are,
\begin{equation}
  \label{eq:sph-discretization-continuity}
  \frac{\tilde{d}\rho_a}{dt} = \sum_{b} \; \frac{m_b}{\rho_{b}} \; (
  \rho_{a} \; \tilde{\ten{u}}_{ab} \; + \;
  (\rho \; (\tilde{\ten{u}} \; - \;
  \ten{u}))_{ab}) \; \cdot \nabla_{a} W_{ab},
\end{equation}

\begin{multline}
  \label{eq:sph-discretization-edac}
  \frac{\tilde{d}p_a}{dt} = \sum_{b} \; \frac{m_b}{\rho_{b}} \; \bigg(
  (p_{a} - \rho_{a} c_{s}^2) \; \ten{u}_{ab} \; + \;
  p_{a} \; \tilde{\ten{u}}_{ab} \; - \;
  (p \; (\tilde{\ten{u}} - \ten{u}))_{ab} \; + \; \\
  4 \; \nu_{edac}
  \frac{p_a - p_b}{(\rho_a + \rho_b) (r^2_{ab} + 0.01 h_{ab}^{2})} \ten{r}_{ab}
  \bigg) \; \cdot \nabla_{a} W_{ab}.
\end{multline}
%
Similarly, the discretized momentum equation for fluids is written as,
\begin{multline}
  \label{eq:sph-momentum-fluid}
  \frac{\tilde{d}\ten{u}_{a}}{dt} = - \sum_{b} m_b \bigg[
  \bigg(\frac{p_a}{\rho_a^2} + \frac{p_b}{\rho_b^2}\bigg) \ten{I} -
  \bigg(\frac{\ten{A}_a}{\rho_a^2} + \frac{\ten{A}_b}{\rho_b^2}
  \bigg) \bigg]
  \cdot \nabla_{a} W_{ab} \\
  + \ten{u}_{a} \sum_{b} \frac{m_b}{\rho_{b}} \; \tilde{\ten{u}}_{ab} \cdot
  \nabla_{a} W_{ab} + \sum_{b} m_b \frac{4 \eta \nabla W_{ab}\cdot
    \ten{r}_{ab}}{(\rho_a + \rho_b) (r_{ab}^2 + 0.01 h_{ab}^2)} \ten{u}_{ab} +
  \ten{g}_{a},
\end{multline}
where $\ten{A}_a = \rho_a \ten{u}_a \otimes (\ten{\tilde{u}}_a - \ten{u}_a)$,
$\ten{I}$ is the identity matrix, $\eta$ is the kinematic viscosity of the
% fluid and \textcite{morris1997modeling} formulation is used to discretize the
viscosity term.

We add to the momentum equation an additional artificial viscosity term
% $\Pi_{ab}$~\parencite{monaghan-review:2005} to maintain the stability of the
numerical scheme, given as,
\begin{align}
  \label{eq:mom-av}
  \Pi_{ab} =
  \begin{cases}
\frac{-\alpha h_{ab} \bar{c}_{ab} \phi_{ab}}{\bar{\rho}_{ab}}
  & \ten{u}_{ab}\cdot \ten{r}_{ab} < 0, \\
  0 & \ten{u}_{ab}\cdot \ten{r}_{ab} \ge 0,
\end{cases}
\end{align}
where,
%
\begin{equation}
  \label{eq:av-phiij}
  \phi_{ab} = \frac{\ten{u}_{ab} \cdot \ten{r}_{ab}}{r^2_{ab} + 0.01 h^2_{ab}},
\end{equation}
%
where $\ten{r}_{ab} = \ten{r}_a - \ten{r}_b$, $\ten{u}_{ab} = \ten{u}_a -
\ten{u}_b$, $h_{ab} = (h_a + h_b)/2$, $\bar{\rho}_{ab} = (\rho_a + \rho_b)/2$,
$\bar{c}_{ab} = (c_a + c_b) / 2$, and $\alpha$ is the artificial
viscosity parameter.

%
For solid mechanics the momentum equation is written as,
\begin{equation}
  \label{eq:sph-momentum-solid}
  \frac{\tilde{d}\ten{u}_{a}}{dt} = - \sum_{b} m_b \bigg[
  \bigg(\frac{p_a}{\rho_a^2} + \frac{p_b}{\rho_b^2}\bigg) \ten{I} -
  \bigg(\frac{\teng{\sigma}^{'}_{a}}{\rho_a^2} +
  \frac{\teng{\sigma}^{'}_{b}}{\rho_b^2} + \Pi_{ab} \ten{I} \bigg) \bigg]  \cdot \nabla_{a} W_{ab} +
  \ten{g}_{a},
\end{equation}
we have not considered the correction stress term $\ten{A}$ in momentum
equation of solid mechanics as it has a negligible effect.

In addition to these three equations, the Jaumann stress rate equation is also
solved. In the current work we use the momentum velocity $\ten{u}$ rather than
% $\tilde{\ten{u}}$ as in the GTVF~\parencite{zhang_hu_adams17} in the computation of
gradient of velocity. The SPH discretization of the gradient of velocity is
given as,
\begin{equation}
  \label{eq:sph-vel-grad}
  \nabla \ten{u}_a =
  - \sum_{b} \frac{m_b}{\rho_{b}} (\ten{u}_{a} - \ten{u}_{b}) \otimes (\nabla_{a} W_{ab}),
\end{equation}
where $\otimes$ is the outer product.

The SPH discretization of the modified Jaumann stress rate
\cref{eq:modified-jaumann-stress-rate} is given as,
\begin{multline}
  \label{eq:sph-modified-jaumann-stress}
  \frac{\tilde{d}\teng{\sigma}^{'}_{a}}{dt} = 2G (\dot{\teng{\epsilon}}_{a} -
  \frac{1}{3} \dot{\teng{\epsilon}}_{a} \ten{I}) + \teng{\sigma}^{'}_{a}
  \teng{\Omega}_{a}^{T} +
  \teng{\Omega}_{a} \teng{\sigma}^{'}_{a} + \\
  + \sum_{b} \; \frac{m_b}{\rho_{b}} \; (\teng{\sigma}^{'} \otimes (\tilde{\ten{u}} -
  \ten{u}))_{ab} \; \cdot \nabla_{a} W_{ab}
  + \teng{\sigma}^{'}_{a} \sum_{b} \; \frac{m_b}{\rho_{b}} \;
  (\tilde{\ten{u}} - \ten{u})_{ab} \; \cdot \nabla_{a} W_{ab}.
\end{multline}

\FloatBarrier%
\section{Rigid body modeling}
\label{sec:conclusions}

\FloatBarrier%
\section{Numerical Methodology}
\label{sec:rfc:rbd}
% The rigid body is discretized into particles with equal spacing each particle
% with mass $m_i$ and density $\rho_i$. Rigid body has a total 6 degrees of
% freedom (DOF), divided into $3$ translational and $3$ rotational.
The equations governing the dynamics of a rigid body are, balance of linear and
angular momentum given by,
\begin{equation}
  \label{eq:rfc:balance_linear_mom}
  \frac{d \; (M \ten{v}_{cm})}{d t} = \sum_i \ten{F}_i,
\end{equation}
\begin{equation}
  \label{eq:rfc:balance_angular_mom}
  \frac{d \ten{L}}{d t} = \teng{\tau}_{cm},
\end{equation}
where $M$, $\ten{v}_{cm}$ are the mass and velocity of the center of mass of the rigid body.
$\ten{F}_i, \teng{\tau}_{cm}, \ten{L} $ are force acting at point $i$, torque and
angular momentum about the center of mass of the rigid body. In the current
case, force acting on the particle $i$, $\ten{F}_i$, is due to the interaction
with the other bodies and with the fluid particles, and any other body forces.
The torque $\teng{\tau}_{cm}$ and angular momentum $\ten{L}$ are computed as,
\begin{equation}
  \label{eq:rfc:torque}
 \teng{\tau}_{cm} = \sum_i \ten{F}_i \times (\ten{x}_{cm} - \ten{x}_{i}),
\end{equation}
\begin{equation}
  \label{eq:rfc:moi}
  \teng{L} =
  \sum_i \; \ten{r}_i \times \; (\teng{\omega} \times \ten{r}_i)
  = \sum_i \; m_i \; [(\ten{r}_i \cdot \ten{r}_i) \ten{I} - \ten{r}_i \otimes \ten{r}_i].
\end{equation}
Here $\ten{x}_{cm}$ and $\omega$ are the position of the center of mass and
angular velocity of the rigid body. $m_i$, $\ten{x}_{i}$, $\ten{r}_i$ are the
mass, position of particle, and position of particle $i$ with respect to vector
center of mass.

% \begin{figure}[!htpb]
%   \centering
%   \includegraphics[width=0.7\textwidth]{images/rfc/images/rigid_body/rigid_body}
%   \caption{Body frame and local frame description of rigid body}
%   \label{fig:gloabl_body_frame_rb}
% \end{figure}
We use two coordinate frames to capture the dynamics of the rigid body, a
global frame and a body frame as shown in
\cref{fig:gloabl_body_frame_rb}. The body fixed frame, which moves with
rigid body is always located at the center of mass ($\ten{x}_{cm}$). The
state of the rigid body at a given time ($t$) can be described using position
($\ten{x}_{cm}$) and velocity ($\ten{v}_{cm}$) of the center of mass, a
rotation matrix($\ten{R}$) to represent the orientation of the rigid body with
respect to the global frame, and angular velocity($\teng{\omega}$). The center
of mass is computed as
\begin{equation}
  \label{eq:rfc:center_of_mass}
  \ten{x}_{cm} = \frac{\sum_i m_i \; \ten{x}_{i} }{\sum_i m_i }.
\end{equation}
The position of the discretized particle ($i$) in
\cref{fig:gloabl_body_frame_rb} belonging to the rigid body at time $t$ can be
computed as,
\begin{equation}
  \label{eq:rfc:rb_particle_pos_update}
  \ten{x}_i = \ten{x}_{cm} + \ten{r}_{i},
\end{equation}
with
\begin{equation}
  \label{eq:rfc:rb_particle_pos_update}
  \ten{r}_i = \ten{R} \overline{\ten{r}}_{i}.
\end{equation}
Here $\overline{\ten{r}}_{i}$ is the position of the particle $i$ about the body
frame axis and remains constant through out the simulation. The rotation matrix
$\ten{R}$ is used to bring the body frame position vector to the global frame
$\ten{O}$. Similarly the velocity vector is computed as,
\begin{equation}
  \label{eq:rfc:rb_particle_vel_update}
  \ten{v}_i = \ten{v}_{cm} + \teng{\omega} \times \ten{r}_{i}.
\end{equation}

@@

% We model the fluid with the CTVF \parencite{adepu2021corrected} scheme
developed in \cref{chap:ctvf}. Similar to the fluid modeling in
\cref{chap:fsi}, we modify the momentum equation to include the force acting due
to the interaction with a solid body. The discretized momentum equation of the
fluid particle including the interaction force is given as
\begin{multline}
  \label{eq:rfc:sph-momentum-fluid}
  \frac{\tilde{d}\ten{u}_{a}}{dt} = - \sum_{b} m_b \bigg[
  \bigg(\frac{p_a}{\rho_a^2} + \frac{p_b}{\rho_b^2}\bigg) \ten{I} -
  \bigg(\frac{\ten{A}_a}{\rho_a^2} + \frac{\ten{A}_b}{\rho_b^2}
  \bigg) \bigg]
  \cdot \nabla_{a} W_{ab} \\
  + \ten{u}_{a} \sum_{b} \frac{m_b}{\rho_{b}} \; \tilde{\ten{u}}_{ab} \cdot
  \nabla_{a} W_{ab} + \sum_{b} m_b \frac{4 \eta \nabla W_{ab}\cdot
    \ten{r}_{ab}}{(\rho_a + \rho_b) (r_{ab}^2 + 0.01 h_{ab}^2)} \ten{u}_{ab} +
  \ten{g}_{a} + \frac{\ten{F}^a_{\text{Fl}}}{m_a}.
\end{multline}
The particles are transported with a transport velocity as described in
\cref{chap:ctvf,chap:fsi}. The boundaries are handled using the dummy particle
approach as described in \cref{chap:ctvf}.

\subsection{Time Integration}

>>>>

  \title{Study of mixing behaviour of spherical particles in fluid using stirrer}
  \author[XXX]{Dinesh Adepu\corref{cor1}}
  \ead{d.dinesh@surrey.ac.uk}
  \author[University of Surrey]{Chuan Yu Wu}
  \ead{XXX}
\address[xxx]{xxx}

@@

\begin{abstract}
  Particle dispersion is a fundamental physical phenomenon encountered in
  various industries where powder and fluid interactions are
  pivotal. Particularly, processes involving the mixing of powder with fluid
  using a stirrer are commonplace. While experimental investigation of such
  processes is often impractical due to their highly nonlinear nature,
  numerical methods offer a viable alternative. Among these, Smoothed Particle
  Hydrodynamics coupled with Discrete Element Method (SPH-DEM) stands out for
  its ability to handle free surface fluid dynamics and model the complex
  physics inherent in dispersion processes.

  In this study, we initially develop and validate both the DEM and SPH
  solvers. Validation includes testing against fundamental benchmarks such as
  two-particle head-on collisions and particle-wall impacts for the DEM
  solver, and Poiseuille flow for the fluid solver. Additionally, we validate
  the rigid-fluid model through simulations of a single cylindrical body
  entering a steady tank and the settling of two tandem cylinders in a
  fluid-filled tank.

  With confidence in the fidelity of our solver established, we delve into
  investigating the mixing behavior of spherical particles dispersed in
  water. Our study encompasses various scenarios, including particles with
  densities higher or lower than the surrounding fluid, as well as particles
  with varying radii. We also examine combinations of these factors.
  \todoin{What factor have you considered to quantify the mixing
    behaviour.}

  \todoin{Write the findings}
\end{abstract}

@@

\FloatBarrier%
\section{Introduction}

\begin{figure}[!htpb]
  \centering
  \includegraphics[width=0.4\textwidth]{images/hs_tank_fluid_with_particles}
  \caption{Body frame and local frame description of rigid body}
  \label{fig:gloabl_body_frame_rb}
\end{figure}

\begin{itemize}
\item Fluid modeling using smoothed particle hydrodynamics
  (\cref{sec:fluid--modeling})
\item Interaction among the discrete spherical rigid particle units by
  discrete element method (\cref{sec:rbd})
\end{itemize}

While the interaction among the rigid spherical particle and fluid particles
are modeled by sampling the spherical particles into dummy SPH particles. This
is explained in \cref{sec:rfc}.

% \FloatBarrier%
% \section{Fluid modeling}

% \FloatBarrier%
% \subsection{Boundary conditions}

% \FloatBarrier%
% \section{Solid phase modeling}

% \FloatBarrier%
% \subsection{Discrete element method}

% \FloatBarrier%
% \subsection{Rigid body equations}

% \FloatBarrier%
% \section{Rigid fluid coupling}

% \FloatBarrier%
% \section{Results}

\FloatBarrier%
\subsection{Fluid modeling}
\label{sec:fluid--modeling}

In the current work, we follow a weakly-compressible SPH approach to model the
fluid. The SPH discretization of the continuity
equation~\cref{eq:sph-discretization-continuity} and the momentum equation
~\cref{eq:sph-momentum-fluid} respectively are,
\begin{equation}
  \label{eq:sph-discretization-continuity}
  \frac{{d}\rho_a}{dt} = \sum_{b} \; \frac{m_b}{\rho_{b}} \;
  \rho_{a} \; {\ten{u}}_{ab} \; \cdot \nabla_{a} W_{ab},
\end{equation}

%
Similarly, the discretized momentum equation for fluids is written as,
\begin{multline}
  \label{eq:sph-momentum-fluid}
  \frac{{d}\ten{u}_{a}}{dt} = - \sum_{b} m_b
  \bigg(\frac{p_a}{\rho_a^2} + \frac{p_b}{\rho_b^2}\bigg)
  \nabla_{a} W_{ab}
 \;+\;
  \sum_{b} m_b \frac{4 \eta \nabla W_{ab}\cdot
    \ten{r}_{ab}}{(\rho_a + \rho_b) (r_{ab}^2 + 0.01 h_{ab}^2)} \ten{u}_{ab}  \;+\;
  \ten{g}_{a} + \text{force due to rb},
\end{multline}
where $\ten{I}$ is the identity matrix, $\eta$ is the kinematic viscosity of the
% fluid and \textcite{morris1997modeling} formulation is used to discretize the
viscosity term.

We add to the momentum equation an additional artificial viscosity term
% $\Pi_{ab}$~\parencite{monaghan-review:2005} to maintain the stability of the
numerical scheme, given as,
\begin{align}
  \label{eq:mom-av}
  \Pi_{ab} =
  \begin{cases}
\frac{-\alpha h_{ab} \bar{c}_{ab} \phi_{ab}}{\bar{\rho}_{ab}}
  & \ten{u}_{ab}\cdot \ten{r}_{ab} < 0, \\
  0 & \ten{u}_{ab}\cdot \ten{r}_{ab} \ge 0,
\end{cases}
\end{align}
where,
%
\begin{equation}
  \label{eq:av-phiij}
  \phi_{ab} = \frac{\ten{u}_{ab} \cdot \ten{r}_{ab}}{r^2_{ab} + 0.01 h^2_{ab}},
\end{equation}
%
where $\ten{r}_{ab} = \ten{r}_a - \ten{r}_b$,
$\ten{u}_{ab} = \ten{u}_a - \ten{u}_b$, $h_{ab} = (h_a + h_b)/2$,
$\bar{\rho}_{ab} = (\rho_a + \rho_b)/2$, $\bar{c}_{ab} = (c_a + c_b) / 2$, and
$\alpha$ is the artificial viscosity parameter.  The pressure $p_a$ is evaluated
using an equation of state:
\begin{equation}
\label{eqn:sph-eos}
  p_a = K \bigg(\frac{\rho_a}{\rho_{0}} - 1 \bigg).
\end{equation}
Where, $K=\rho_0 \, c_0^2$ is bulk modulus of the body, with
$c_0=10 \times V_{\text{max}}$ is speed of sound, while $\rho_0$ as the
initial density of the particles.

\FloatBarrier%
\subsection{Boundary conditions}
\label{sec:boundary_conditions}

\FloatBarrier%
\section{Rigid body dynamics}
\label{sec:rbd}
% The rigid body is discretized into particles with equal spacing each particle
% with mass $m_i$ and density $\rho_i$. Rigid body has a total 6 degrees of
% freedom (DOF), divided into $3$ translational and $3$ rotational.
The equations governing the dynamics of a rigid body are, balance of linear and
angular momentum given by,
\begin{equation}
  \label{eq:rfc:balance_linear_mom}
  \frac{d \; (M \ten{v}_{cm})}{d t} = \sum_i \ten{F}_i,
\end{equation}
\begin{equation}
  \label{eq:rfc:balance_angular_mom}
  \frac{d \ten{L}}{d t} = \teng{\tau}_{cm},
\end{equation}
where $M$, $\ten{v}_{cm}$ are the mass and velocity of the center of mass of the rigid body.
$\ten{F}_i, \teng{\tau}_{cm}, \ten{L} $ are force acting at point $i$, torque and
angular momentum about the center of mass of the rigid body. In the current
case, force acting on the particle $i$, $\ten{F}_i$, is due to the interaction
with the other bodies and with the fluid particles, and any other body forces.
The torque $\teng{\tau}_{cm}$ and angular momentum $\ten{L}$ are computed as,
\begin{equation}
  \label{eq:rfc:torque}
 \teng{\tau}_{cm} = \sum_i \ten{F}_i \times (\ten{x}_{cm} - \ten{x}_{i}),
\end{equation}
\begin{equation}
  \label{eq:rfc:moi}
  \teng{L} =
  \sum_i \; \ten{r}_i \times \; (\teng{\omega} \times \ten{r}_i)
  = \sum_i \; m_i \; [(\ten{r}_i \cdot \ten{r}_i) \ten{I} - \ten{r}_i \otimes \ten{r}_i].
\end{equation}
Here $\ten{x}_{cm}$ and $\omega$ are the position of the center of mass and
angular velocity of the rigid body. $m_i$, $\ten{x}_{i}$, $\ten{r}_i$ are the
mass, position of particle, and position of particle $i$ with respect to vector
center of mass.

\begin{figure}[!htpb]
  \centering
  \includegraphics[width=0.7\textwidth]{images/rigid_body/rigid_body}
  \caption{Body frame and local frame description of rigid body}
  \label{fig:gloabl_body_frame_rb}
\end{figure}
We use two coordinate frames to capture the dynamics of the rigid body, a
global frame and a body frame as shown in
\cref{fig:gloabl_body_frame_rb}. The body fixed frame, which moves with
rigid body is always located at the center of mass ($\ten{x}_{cm}$). The
state of the rigid body at a given time ($t$) can be described using position
($\ten{x}_{cm}$) and velocity ($\ten{v}_{cm}$) of the center of mass, a
rotation matrix($\ten{R}$) to represent the orientation of the rigid body with
respect to the global frame, and angular velocity($\teng{\omega}$). The center
of mass is computed as
\begin{equation}
  \label{eq:rfc:center_of_mass}
  \ten{x}_{cm} = \frac{\sum_i m_i \; \ten{x}_{i} }{\sum_i m_i }.
\end{equation}
The position of the discretized particle ($i$) in
\cref{fig:gloabl_body_frame_rb} belonging to the rigid body at time $t$ can be
computed as,
\begin{equation}
  \label{eq:rfc:rb_particle_pos_update}
  \ten{x}_i = \ten{x}_{cm} + \ten{r}_{i},
\end{equation}
with
\begin{equation}
  \label{eq:rfc:rb_particle_pos_update}
  \ten{r}_i = \ten{R} \overline{\ten{r}}_{i}.
\end{equation}
Here $\overline{\ten{r}}_{i}$ is the position of the particle $i$ about the body
frame axis and remains constant through out the simulation. The rotation matrix
$\ten{R}$ is used to bring the body frame position vector to the global frame
$\ten{O}$. Similarly the velocity vector is computed as,
\begin{equation}
  \label{eq:rfc:rb_particle_vel_update}
  \ten{v}_i = \ten{v}_{cm} + \teng{\omega} \times \ten{r}_{i}.
\end{equation}

@@

\FloatBarrier%
\subsection{Contact resolution}
\label{sec:dem}

\begin{figure}[!htpb]
  \centering
  \includegraphics[width=0.7\textwidth]{images/spherical_particles_dem_representation}
  \caption{Body frame and local frame description of rigid body}
  \label{fig:gloabl_body_frame_rb}
\end{figure}

We resolve the contact among the spherical particles using discrete element
method. In the current work we have utilized the contact force model proposed
by \citet{mohseni2021particle}. The force acting on a particle $a$ due to the
interaction with the particle $b$ can be resolved into a normal and tangential
component. The normal force component is utilised to make sure that the
particle do not penetrate into each other, while the tangential component is
used to model the friction between the interacting solids.  The normal force
($\teng{F}_a^{n}$) on particle $a$ due to the interaction with the particles
$b$ is given as,
\begin{equation}
  \label{eq:contact-algorithm-normal}
  \ten{F}_a^n = k_r \delta_{n^{c}}^{a} \ten{n}_a^{c}.
\end{equation}
Here, the overlap $\delta_{n^{c}}^{a}$ is computed using
\begin{equation}
  \label{eq:cf-overlap}
  \delta_{n^{c}}^{a} = \Delta x - d_a,
\end{equation}
where the normal contact vector $\ten{n}_a^{c}$ is computed using
\begin{equation}
  \label{eq:cf-normal-vector}
  \ten{\hat{n}}_a^{c} = \frac{
    \displaystyle\sum\limits_{b = 1}^{\text{NP}^{b}} \;
    \frac{\ten{r}_{ab}}{r_{ab}}  \frac{m_b}{\rho_b} W_{ab}}
  {
    \displaystyle\sum\limits_{b = 1}^{\text{NP}^{b}} \;
    \frac{m_b}{\rho_b} W_{ab}},
\end{equation}
\begin{equation}
  \label{eq:cf-normal-vector}
  \ten{n}_a^{c} = \frac{\teng{\hat{n}}_a^{c}}{||\teng{\hat{n}}_a^{c}||}.
\end{equation}
With $\Delta x$ being the initial spacing between the particles, $k_r$ is the normal
spring stiffness coefficient. Note that while computing the overlap of
particle $a$ with the body $B$ we have computed an effective overlap, rather
than per particle interaction. This effectively is able to model the
interaction between non smooth surfaces in contrast with particle-particle
force computation.

\subsection{Tangential force computation}
\label{sec:tangential-force-computation}
We associate a tangential spring attached to particle $i$ and body $B$ to
compute the tangential force, which initially has a magnitude of zero
($|\Delta \textit{\textbf{l}}_a|=0$). And the tangential spring is activated
when the particle comes into contact with body $B$. The tangential force is
history-dependent. The contact friction force is proportional to the
tangential spring displacement, which is integrated over the contact time as
\begin{equation}
  \label{eq:tangential-force}
  \ten{F}_{a}^{t^{n+1}} =
  -k_f \Delta \textit{\textbf{l}}_a^{\,n + 1} =
  -k_f \big[\big(\Delta {\textit{\textbf{l}}}_a^{\,n} \
  + \ten{v}_{ab}^{n + 1} \Delta t\big) \cdot \ten{t}_a^{n + 1} \big] \
  \ten{t}_a^{n + 1},
\end{equation}
where $\Delta t$ is the time step, $\ten{v}_{ab} = \ten{v}_{a} - \ten{v}_b$ is
the relative velocity of the primary particle $a$ with respect to the
closest secondary particle $b$, and $k_f$ is the tangential spring stiffness
coefficient. The tangential unit vector is computed by,
\begin{equation}
  \label{eq:tangential-vect}
  \ten{t}_a = \frac{\ten{v}_{ab} - (\ten{v}_{ab} \cdot \ten{n}_a^{c}) \ten{n}_a^{c}}{|\ten{v}_{ab} - (\ten{v}_{ab} \cdot \ten{n}_a^{c}) \ten{n}_a^{c}|}.
\end{equation}
